---
title: "Test a Perceptual Phenomenon"
output: pdf_document
---

```{r setup, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
setwd("~/Udacity/Project 5 - Statistics")
library(ggplot2)
library(gridExtra)
library(nortest)
```

```{r}
stroop <- read.csv('stroopdata.csv')
```

###What is our independent variable?  What is our dependent variable?

The independent variable in this dataset is the type of condition (i.e. the
congruent words condition and the incongruent words condition).  The dependent
variable is the time taken for participants to go through the list for each 
condition.

###What is an appropriate set of hypotheses for this task?  What kind of \
statistical test do you expect to perform?  Justify your choices.

The set of hypotheses for this test will be:

$H_{0}$: $\mu_{I}$ - $\mu_{C}$ $=$ 0

(There is no significant difference between the population mean completion time
for the incongruent words condition ($\mu_{I}$) and the population mean
completion time for the congruent words condition ($\mu_{C}$)).

$H_{A}$: $\mu_{I}$ - $\mu_{C}$ $>$ 0

(The difference between the population mean completion time for the incongruent
words condition and the population mean completion time for the congruent words
condition is greater than zero).


I decided the appropriate statistical test to perform will be the paired sample
t-test since:

* We are working with dependent samples since we are comparing two sets of times
for tests taken by the same subject.  

* We do not know population standard deviations.  

* Our sample size is less than 30

That said, one crucial assumption we need to make in order for us to use the 
paired t-test is that the data we sample from comes from a normal distribution.
We will check this assumption later in the report by visaulizing our sample 
data.

###Report some descriptive statistics regarding this dataset. Include at least\
one measure of central tendency and at least one measure of variability.

```{r}
summary(stroop$Congruent)
sd(stroop$Congruent)
```

Here are some descriptive statistics for the congruent condition data.  We can 
see that the mean time is 14.05 with a standard deviation of 3.559.  That said,
what will the incongruent condition data look like?

```{r}
summary(stroop$Incongruent)
sd(stroop$Incongruent)
```

The mean time is 22.02 while the standard deviation is 4.797.  Let's get another
look at the distribution of the data with visualizations.

### Provide one or two visualizations that show the distribution of the sample \
data. Write one or two sentences noting what you observe about the plot or \
plots.

```{r}
p1 <- qplot(x = stroop$Congruent, binwidth = 24^(1/3), color = I('black'),
            fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(5, 25, 5)) +
  xlab('Time to Completion') +
  ylab('Number of Participants') +
  ggtitle('Histogram of Times for the Congruent Condition')

p2 <- qplot(x = stroop$Incongruent, binwidth = 24^(1/3), color = I('black'),
            fill = I('#5760AB')) +
  scale_x_continuous(breaks = seq(15, 35, 5)) +
  xlab('Time to Completion') +
  ylab('Number of Participants') +
  ggtitle('Histogram of Times for the Incongruent Condition')

grid.arrange(p1, p2)
```

Looking at the histogram for the congruent condition data, it appears as though
the data is most likely normally distributed.  On the other hand, the histogram
for the incongruent condition data appears as though it may have some outlier
participants which may affect the normality of the data.

```{r}
p3 <- ggplot(data = stroop, aes(sample = Congruent)) +
  stat_qq()
p4 <- ggplot(data = stroop, aes(sample = Incongruent)) +
  stat_qq()

grid.arrange(p3, p4)

ad.test(stroop$Congruent)
ad.test(stroop$Incongruent)
```

If we look at normal probability plot for the congruent data, the points appears
to fall a relatively straight line and is thus most likely normally distributed.
If we perform an Anderson-Darling test of normality on the data, where 
$\alpha$ = 0.05, we can see that the p-value for the test is 0.8432.  Since the 
p-value is above 0.05, there is not enough evidence to conclude that the data 
is not normally distributed.

However, If we look at the probability plot for the incongruent data, two points
on the right do not fall on the with the rest of the data, which suggests the 
data may not be normally distributed.  Performing an Anderson-Darling test, with
$\alpha$ = 0.05, the p-value is 0.0067, which is less than 0.05, therefore there
is evidence that this data is not normally distributed.  This poses a problem
for our paired t-test, since a requirement of performing the test is that the 
data we use is normally distributed or if it is not normally distributed, we are
using a very large sample.  Since we cannot fulfill this requirement, typically
this means that we cannot perform the t-test and should perform some other kind
of test.  That being said, for the sake of completing this project I will go 
ahead and perform the paired t-test.

###Now, perform the statistical test and report your results.  What is your \
confidence level and your critical statistic value?  Do you reject the null \
hypothesis or fail to reject it?  Come to a conclusion in terms of the \
experiment task.  Did the results match up with your expectations?

The confidence level I decided to use an $\alpha$ = 0.05 meaning our confidence
level will be 0.95.  The test statistic I came up with is 8.0207.  Using a 
t-table, the critical statistic value, with 23 degrees of freedom, is 1.7139.  
So, since the test statistic (8.0207) is greater than the critical value 
(1.7139) we should reject the null hypothesis.  We can conclude that there is 
evidence that the difference between the population mean completion time for the incongruent condition and the congruent condition is greater than 0.

##Conclusion

From these results it appears as though participants in the Stroop task, take
significantly longer when it comes to completing the list of incongruent words.
These findings happened to match up with my expectations I personally took 
longer to complete the incongruent condition when performing the Stroop task.

##References

https://www.r-bloggers.com/normality-tests-for-continuous-data/

http://ggplot2.tidyverse.org/reference/geom_qq.html

http://www.statisticshowto.com/when-to-use-a-t-score-vs-z-score/
